### 2017-04-27 DLM/writeup.tex
Added the SVM model

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{url}
\usepackage{wasysym}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\setcounter{MaxMatrixCols}{30}
\providecommand{\U}[1]{\protect\rule{.1in}{.1in}}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\numberwithin{equation}{section}

@@

Many models of interest contain a set of global parameters $\boldsymbol{\theta}$ and observation specific latent variables $\textbf{x}_{1:T}$ such that $\textbf{x}_t$ is assumed to follow a Markovian structure. In this case $y_{T+S+h}$ depends on the latent variables $\textbf{x}_{T+1:T+S+h}$, and the forecast uncertainty in $y_{T+S+h}$ can be reduced by first incorporating information contained in realisations of $y_{T+1:T+S}$ to form the posterior densities of $\textbf{x}_{T+1:T+S+h} | \textbf{x}_{1:T}, \boldsymbol{\theta}, y_{1:T+S}$. With certain restrictions on the model, the additional information can be easily included via filtering techniques such as the Kalman filter, however in many cases these restrictions are unsuitable and a computationally costly re-estimation of posterior densities of parameters and latent states must be undertaken.

One of the most common computational techniques for models with latent states is Markov Chain Monte Carlo (MCMC), which involves iterative sampling from a Markov Chain designed to converge to the true posterior distribution. MCMC can have slow convergence rates and generally does not admit parallel processing so that updating the forecast distribution of $y_{T+S+h} | y_{1:T}$ to $y_{T+S+h} | y_{1:T+S}$ after observation of $y_{T+1}, \dots, y_{T+S}$ may not be available before $y_{T+S+h}$ is observed.

This paper explores the use of Variational Bayes to approximate the posterior distribution of several common time-series models with dependent latent states, and compares the increased speed of computation with the decreased statistical accuracy associated with using an approximation to the posterior distribution. We compare several different functional forms of the approximating distribution that are commonly used in the machine learning literature and explore the use of MCMC samples to inform the design of the functional form.

@@

Variational Bayes posits a divergence function between the true posterior distribution $p(\boldsymbol{\theta}, \textbf{x}_{1:T} | y_{1:T}$ and some approximating distribution $q(\boldsymbol{\theta}, \textbf{x} | \boldsymbol{\lambda})$, choosing the parameters $\boldsymbol{\lambda}$ for a given functional form $q$ that minimises the divergence function.

We follow the traditional approach, where the divergence function is the Kullback-Leibler (KL) divergence (Kullback and Leibler, 1951) from $q(\boldsymbol{\theta}, \textbf{x}_{1:T}| \boldsymbol{\lambda})$ to the true posterior $p(\boldsymbol{\theta}, \textbf{x}_{1:T} | y_{1:T})$. The KL divergence is defined by
\begin{equation}
\label{KL-def}
KL[q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda})\hspace{.1cm}||\hspace{.1cm}p(\boldsymbol{\theta}, \textbf{x}_{1:T} | y_{1:T})] = \int q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}) \ln \left( \frac{q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda})}{p(\boldsymbol{\theta}, \textbf{x}_{1:T} |y_{1:T})}\right) d\boldsymbol{\theta} d\textbf{x}_{1:T}
\end{equation}
and can alternatively be expressed as
\begin{equation}
\label{KL-ELBO}
KL[q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda})\hspace{.1cm}||\hspace{.1cm}p(\boldsymbol{\theta}, \textbf{x}_{1:T} | y_{1:T})] = \ln(p(y_{1:T})) - \mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}), y_{1:T})
\end{equation}
where $\mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}), y_{1:T})$ is referred to as the Evidence Lower Bound (ELBO), as it provides a lower bound on the unknown constant $\ln(p(y_{1:T}))$.  $\mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}), y_{1:T})$ is defined by
\begin{equation}
\label{ELBO}
\mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}), y_{1:T}) = \int q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}) \ln \left( \frac{p(y_{1:T},\boldsymbol{\theta}, \textbf{x}_{1:T})}{q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda})} \right) d\boldsymbol{\theta}d\textbf{x}_{1:T},
\end{equation}
and as $\ln(p(y_{1:T}))$ is constant with respect to $\boldsymbol{\lambda}$, maximising (\ref{ELBO}) with respect to $\boldsymbol{\lambda}$ is equivalent to minimising (\ref{KL-def}). Equation (\ref{ELBO}) can be maximised using a gradient ascent approach, where we iteratively apply the following update step until (\ref{ELBO}) converges within some pre-specified tolerance:
\begin{equation}
\label{GradAscent}
\boldsymbol{\lambda}^{(m+1)} = \boldsymbol{\lambda}^{(m)} + \rho^{(m)} \nabla_{\boldsymbol{\lambda}} \mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}).
\end{equation}
$\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ is the vector of partial derivatives of $\mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ with respect to each element of $\boldsymbol{\lambda}$ evaluated at $\boldsymbol{\lambda}^{(m)}$. This update requires some initial values $\boldsymbol{\lambda}^{(0)}$ and a sequence $\rho^{(m)}, m = 1, 2, \dots$ known as the learning rate. If $\rho^{(m)}$ is chosen to satisfy the following conditions the algorithm is guaranteed to converge to a local maximum (Robbins and Munro, 1951).
\begin{align}
&\sum_{m=1}^{\infty} \rho^{(m)} =  \infty \\
&\sum_{m=1}^{\infty} (\rho^{(m)})^2 <  \infty.
\end{align}

Ranganath, Gerrish and Blei (2014) showed that a Monte Carlo estimate of the derivative of the ELBO can be given by
\begin{equation}
\label{ScoreDeriv}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\boldsymbol{\lambda}} [\ln(q(\boldsymbol{\theta}_s, \textbf{x}_{1:T, s} | \boldsymbol{\lambda}^{(m)}))] \ln \left(\frac{p(y_{1:T}, \boldsymbol{\theta}_s, \textbf{x}_{1:T, s})}{q(\boldsymbol{\theta}_s, \textbf{x}_{1:T, s} | \boldsymbol{\lambda}^{(m)})} \right) 
\end{equation}
where $s = 1, \dots, S$ indicates simulations from $q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}^{(m)})$. 
The terms in the sum in (\ref{ScoreDeriv}) can have large variances, and in practice a large value of $S$ is required to ensure a precise estimate of the gradient of the ELBO is obtained, slowing computation.

The variance can be reduced by the reparameterisation trick of Kingma and Welling (2014), which introduces parameter free noise variables $\boldsymbol{\epsilon}$ and a differentiable transform $f$ such that
\begin{equation}
\label{Reparam}
\boldsymbol{\theta}, \textbf{x}_{1:T} = f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}).
\end{equation}
Kingma and Welling (2014) show that an $f$ exists to transform $\boldsymbol{\epsilon}$ to any continuous random variable, with examples including a location-scale transform of a standard normal noise variable and an inverse-CDF transform of a uniform noise variable. Transforming the variables using
\begin{equation}
\label{changevar}
\int q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}) g(\boldsymbol{\theta}, \textbf{x}_{1:T})  d\boldsymbol{\theta}d\textbf{x}_{1:T} = \int q(\boldsymbol{\epsilon}) g ( f (\epsilon, \boldsymbol{\lambda})) d\boldsymbol{\epsilon})
\end{equation}
and noting that
\begin{equation}
\label{ReparamDist}
\ln(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda})) = \ln(q(\boldsymbol{\epsilon})) - \ln(|J|),
\end{equation}
where $J$ is the Jacobian Matrix of the transformation $f$, allows (\ref{ELBO}) to be rewritten as
\begin{align}
\label{reparamELBO}
\mathcal{L}(q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda}), y_{1:T}) &=  \int  q(\boldsymbol{\epsilon}) \ln \left( \frac{p(y_{1:T}, f (\epsilon, \boldsymbol{\lambda}))}{q( f (\epsilon, \boldsymbol{\lambda}) | \boldsymbol{\lambda})} \right) d\boldsymbol{\epsilon} \nonumber \\
&= \int  q(\boldsymbol{\epsilon}) \big( \ln (p(y_{1:T}, f (\epsilon, \boldsymbol{\lambda})) - \ln(q(\boldsymbol{\epsilon})) + \ln(|J|) \big) d\boldsymbol{\epsilon} \nonumber \\
&= \mathcal{L}(\boldsymbol{\lambda}, y_{1:T}).
\end{align}

The derivative of (\ref{reparamELBO}) can be estimated by
\begin{equation}
\label{ReparamDeriv}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(\boldsymbol{\lambda}^{(m)}, y_{1:T}) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\boldsymbol{\lambda}} \ln (p(y_{1:T}, f(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_s))) + \nabla_{\boldsymbol{\lambda}}\ln(|J(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_s)|),
\end{equation}
where simulations of $\boldsymbol{\theta}$ and $\textbf{x}_{1:t}$ are replaced by simulations of $\boldsymbol{\epsilon}$ from $q(\boldsymbol{\epsilon})$. The variance of the terms in the sum of this estimator is often orders of magnitude lower than the estimator in (\ref{ScoreDeriv}); often $S = 1$ provides a precise enough estimate of the gradient for fast convergence.

@@

@@

We transform $\sigma^2_y$ and $\sigma^2_x$ so that $\boldsymbol{\theta} = \{\gamma, \phi, \ln(\sigma^2_y), \ln(\sigma^2_x) \} \in \mathbb{R}^4$ and consider four different functional forms for the approximating distribution $q(\boldsymbol{\theta}, \textbf{x}_{1:T} | \boldsymbol{\lambda})$:
\begin{enumerate}
\item $q(\boldsymbol{\theta}, \textbf{x} | \boldsymbol{\lambda}) = \mathcal{N}(\boldsymbol{\psi}, \textbf{V})$, where $\textbf{V}$ is a diagonal matrix.
\item $q(\boldsymbol{\theta}, \textbf{x} | \boldsymbol{\lambda}) = \mathcal{N}(\boldsymbol{\psi}, \boldsymbol{\Sigma})$, where $\boldsymbol{\Sigma}$ allows non-zero off-diagonal terms.
\item $q(\boldsymbol{\theta}, \textbf{x} | \boldsymbol{\lambda})$ is an Inverse Autoregressive flow (Kingma, Salimans, Jozefowicz, Chen, Sutskever and Welling, 2016).
\item  $q(\boldsymbol{\theta}, \textbf{x} | \boldsymbol{\lambda}) = \prod \left(q(\theta_i) q(x_{it})\right) c(\boldsymbol{\theta}, \textbf{x})$ where the form of the copula $c$ and marginals are informed by an out of date MCMC sample.
\end{enumerate}
The first three functional forms are transformation of standard normal noise variables while the fourth is a transformation of uniform noise variables.

As additional data $y_{T+1:T+J-1}$ are observed, we can incorporate this information to improve forecasts of the future value of $y_{T+J}$. We use two MCMC based approaches for updating this forecast,
\begin{enumerate}
\item Use the Kalman filter to update $\textbf{x}_{T+1:T+J}$ as $y_{T+1:T+J}$ is observed using $\boldsymbol{\theta}$ from $p(\boldsymbol{\theta}, \textbf{x}_{1:T} | y_{1:T})$. This method is computationally simple, but does not allow the posterior distribution of $\boldsymbol{\theta}$ to be conditioned on new observations.
\item Run the full MCMC for $\boldsymbol{\theta}$ and $\textbf{x}_{1:T+J}$ after observing each additional $y$. This strategy has a high computational cost, but does condition $\boldsymbol{\theta}$ on $y_{T+1:T+J-1}$, and results in the forecast distribution with the least uncertainty.
\end{enumerate}

>>>>

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{tikz}
\usetikzlibrary{bayesnet}
\usepackage{url}
\usepackage{wasysym}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\setcounter{MaxMatrixCols}{30}
\providecommand{\U}[1]{\protect\rule{.1in}{.1in}}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\numberwithin{equation}{section}

@@

Many models of interest contain a set of global parameters $\boldsymbol{\theta}$ and observation specific latent variables $x_{1:T}$ such that $x_t$ is assumed to follow a Markovian structure. In this case $y_{T+S+h}$ depends on the latent variables $x_{T+1:T+S+h}$, and the forecast uncertainty in $y_{T+S+h}$ can be reduced by first incorporating information contained in realisations of $y_{T+1:T+S}$ to form the posterior densities of $x_{T+1:T+S+h} | x_{1:T}, \boldsymbol{\theta}, y_{1:T+S}$. With certain restrictions on the model, the additional information can be easily included via filtering techniques such as the Kalman filter, however in many cases these restrictions are unsuitable and a computationally costly re-estimation of posterior densities of parameters and latent states must be undertaken.

One of the most common computational techniques for models with latent states is Markov Chain Monte Carlo (MCMC), which involves iterative sampling from a Markov Chain designed to converge to the true posterior distribution. MCMC can have slow convergence rates and generally does not admit parallel processing so that updating the forecast distribution of $y_{T+S+h} | y_{1:T}$ to $y_{T+S+h} | y_{1:T+S}$ after observation of $y_{T+1}, \dots, y_{T+S}$ may not be available before $y_{T+S+h}$ is observed. In this paper we assume that an MCMC will converging before time period $T + S + h$ will be required to begin, at the latest, by time $T$ and hence $y_{T}$ is the last observation that can be included in the MCMC.

Increasing the speed of estimation will allow more data up to time $T + S$ to be included in the forecast distribution, and this paper explores the use of Variational Bayes to approximate the posterior distribution of several common time-series models with dependent latent states. The the increase in computation speed and resulting increase in observations included in a forecast is contrasted with the decreased statistical accuracy associated with using an approximation to the posterior distribution. We compare several different functional forms of the approximating distribution that are commonly used in the machine learning literature and explore the use of previous MCMC samples to inform the design of the functional form.

@@

Variational Bayes posits a divergence function between the true posterior distribution $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T}$ and some approximating distribution $q(\boldsymbol{\theta}, x | \boldsymbol{\lambda}$, choosing the parameters $\boldsymbol{\lambda}$ for a given functional form $q$ that minimises the divergence function.

We follow the traditional approach, where the divergence function is the Kullback-Leibler (KL) divergence (Kullback and Leibler, 1951) from $q(\boldsymbol{\theta}, x_{1:T}| \boldsymbol{\lambda})$ to the true posterior $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})$. The KL divergence is defined by
\begin{equation}
\label{KL-def}
KL[q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})\hspace{.1cm}||\hspace{.1cm}p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})] = \int q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}) \ln \left( \frac{q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})}{p(\boldsymbol{\theta}, x_{1:T} |y_{1:T})}\right) d\boldsymbol{\theta} dx_{1:T}
\end{equation}
and can alternatively be expressed as
\begin{equation}
\label{KL-ELBO}
KL[q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})\hspace{.1cm}||\hspace{.1cm}p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})] = \ln(p(y_{1:T})) - \mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}), y_{1:T})
\end{equation}
where $\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}), y_{1:T})$ is referred to as the Evidence Lower Bound (ELBO), as it provides a lower bound on the unknown constant $\ln(p(y_{1:T}))$.  $\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}), y_{1:T})$ is defined by
\begin{equation}
\label{ELBO}
\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}), y_{1:T}) = \int q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}) \ln \left( \frac{p(y_{1:T},\boldsymbol{\theta}, x_{1:T})}{q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})} \right) d\boldsymbol{\theta}dx_{1:T},
\end{equation}
and as $\ln(p(y_{1:T}))$ is constant with respect to $\boldsymbol{\lambda}$, maximising (\ref{ELBO}) with respect to $\boldsymbol{\lambda}$ is equivalent to minimising (\ref{KL-def}).

Equation (\ref{ELBO}) can be maximised using a gradient ascent approach, where we iteratively apply the following update step until (\ref{ELBO}) converges within some pre-specified tolerance:
\begin{equation}
\label{GradAscent}
\boldsymbol{\lambda}^{(m+1)} = \boldsymbol{\lambda}^{(m)} + \rho^{(m)} \nabla_{\boldsymbol{\lambda}} \mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}).
\end{equation}
$\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ is the vector of partial derivatives of $\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ with respect to each element of $\boldsymbol{\lambda}$ evaluated at $\boldsymbol{\lambda}^{(m)}$. This update requires some initial values $\boldsymbol{\lambda}^{(0)}$ and a sequence $\rho^{(m)}, m = 1, 2, \dots$ known as the learning rate. If $\rho^{(m)}$ is chosen to satisfy the following conditions the algorithm is guaranteed to converge to a local maximum (Robbins and Munro, 1951).
\begin{align}
&\sum_{m=1}^{\infty} \rho^{(m)} =  \infty \\
&\sum_{m=1}^{\infty} (\rho^{(m)})^2 <  \infty.
\end{align}

Ranganath, Gerrish and Blei (2014) showed that a Monte Carlo estimate of the derivative of the ELBO can be given by
\begin{equation}
\label{ScoreDeriv}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{N}\sum_{i=1}^{N} \nabla_{\boldsymbol{\lambda}} [\ln(q(\boldsymbol{\theta}_i, x_{1:T, i} | \boldsymbol{\lambda}^{(m)}))] \ln \left(\frac{p(y_{1:T}, \boldsymbol{\theta}_i, x_{1:T, i})}{q(\boldsymbol{\theta}_i, x_{1:T, i} | \boldsymbol{\lambda}^{(m)})} \right) 
\end{equation}
where $i = 1, \dots, N$ indicates simulations from $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}^{(m)})$. 
The terms in the sum in (\ref{ScoreDeriv}) can have large variances, and in practice a large value of $N$ is required to ensure a precise estimate of the gradient of the ELBO is obtained, slowing computation.

The variance can be reduced by the reparameterisation trick of Kingma and Welling (2014), which introduces parameter free noise variables $\boldsymbol{\epsilon}$ and a differentiable transform $f$ such that
\begin{equation}
\label{Reparam}
\boldsymbol{\theta}, x_{1:T} = f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}).
\end{equation}
Kingma and Welling (2014) show that an $f$ exists to transform $\boldsymbol{\epsilon}$ to any continuous random variable, with examples including a location-scale transform of a standard normal noise variable and an inverse-CDF transform of a uniform noise variable. Transforming the variables using
\begin{equation}
\label{changevar}
\int q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}) g(\boldsymbol{\theta}, x_{1:T})  d\boldsymbol{\theta}dx_{1:T} = \int q(\boldsymbol{\epsilon}) g ( f (\epsilon, \boldsymbol{\lambda})) d\boldsymbol{\epsilon})
\end{equation}
and noting that
\begin{equation}
\label{ReparamDist}
\ln(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})) = \ln(q(\boldsymbol{\epsilon})) - \ln(|J|),
\end{equation}
where $J$ is the Jacobian Matrix of the transformation $f$, allows (\ref{ELBO}) to be rewritten as
\begin{align}
\label{reparamELBO}
\mathcal{L}(q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda}), y_{1:T}) &=  \int  q(\boldsymbol{\epsilon}) \ln \left( \frac{p(y_{1:T}, f (\epsilon, \boldsymbol{\lambda}))}{q( f (\epsilon, \boldsymbol{\lambda}) | \boldsymbol{\lambda})} \right) d\boldsymbol{\epsilon} \nonumber \\
&= \int  q(\boldsymbol{\epsilon}) \big( \ln (p(y_{1:T}, f (\epsilon, \boldsymbol{\lambda})) - \ln(q(\boldsymbol{\epsilon})) + \ln(|J|) \big) d\boldsymbol{\epsilon} \nonumber \\
&= \mathcal{L}(\boldsymbol{\lambda}, y_{1:T}).
\end{align}

The derivative of (\ref{reparamELBO}) can be estimated by
\begin{equation}
\label{ReparamDeriv}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(\boldsymbol{\lambda}^{(m)}, y_{1:T}) \approx \frac{1}{N}\sum_{i=1}^{N} \nabla_{\boldsymbol{\lambda}} \ln (p(y_{1:T}, f(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_i))) + \nabla_{\boldsymbol{\lambda}}\ln(|J(\boldsymbol{\lambda}^{(m)}, \boldsymbol{\epsilon}_i)|),
\end{equation}
where simulations of $\boldsymbol{\theta}$ and $x_{1:t}$ are replaced by simulations of $\boldsymbol{\epsilon}$ from $q(\boldsymbol{\epsilon})$. The variance of the terms in the sum of this estimator is often orders of magnitude lower than the estimator in (\ref{ScoreDeriv}); often $N = 1$ provides a precise enough estimate of the gradient for fast convergence.

@@

A key decision in Variational Bayes is the choice of the functional form of the approximating distribution $q$. Much of the literature focuses on a generic form of $q$ that can flexibly be applied to many models. However the time-series nature of our application implies that MCMC samples of $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})$, which provides useful information for constructing the functional form of $q(\boldsymbol{\theta}, x_{1:T+S} | \boldsymbol{\lambda})$ for use in the approximate predictive density, $q(y_{T+S+h} | y_{1:T+S})$. There are two stages to this construction:
\begin{enumerate}
\item Find a well fitting functional form for $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ using the MCMC results.
\item Extrapolate the form of $q$ used for $x_{1:T}$ to $x_{1:T+S}$.
\end{enumerate}
Treating the MCMC samples as 'data', the functional form of $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$ can be approached as a traditional distribution fitting problem. Optimal marginal distributions are selected for each parameter by AIC, while a vine copula modelling approach is adopted for the dependence structure following Tran, Blei and Airoldi (2015). Sklar (1959) proves that any joint probability distribution with $k$ variables can be written as the product of the marginals and a copula function,
\begin{equation}
\label{vc1}
p(\omega_1, \dots, \omega_k) = p(\omega_1) \dots p(\omega_k) c(P(\omega_1), \dots, P(\omega_k))
\end{equation}
where $p(\theta)$ is a pdf, $P(\theta)$ is a cdf, and $c(\cdot)$ is a copula. A vine copula is a structure to factorise this copula, increasing the flexibility given to the modeller, and is defined by a set of trees $\mathcal{V} = \{T_1, \dots, T_{k-1} \}$, where
\begin{enumerate}
\item $T_1$ has $k$ nodes corresponding to the $k$ variables, with $k-1$ edges.
\item $T_t, t = 2, \dots, p-1$ has nodes corresponding to the edges in tree $T_{t-1}$
\item Two nodes in tree $T_t$ are joined by an edge only if the associated edges in $T_{t-1}$ share a node. 
\end{enumerate}
Each edge in the vine corresponds to a bivariate copula conditioned on any variables that are shared between nodes, and the product of each of these $k(k-1)/2$ bivariate copulas forms the complete copula. Figure (\ref{fig:vinecop}) provides an example for a four dimensional $\theta$.
\begin{figure}[h]
\centering
\includegraphics[width=0.4\linewidth,height=\textheight,keepaspectratio]{vines.png}
\vspace{2mm}
\caption{An example of a Vine Copula over four variables. The edges in the top tree represent unconditional bivariate copulas between the parameters on the connected nodes. The second tree edges represent the bivariate copulas for $\theta_1 \theta_3 | \theta_2$ and for $\theta_3 \theta_4 | \theta_2$, the conditioning on $\theta_2$ caused by its appearance in each connected node. The edge in the final tree represents the bivariate copula for $\theta_1, \theta_4 | \theta_2, \theta_3$, as both $\theta_2$ and $\theta_3$ appear in the connected nodes.}
\label{fig:vinecop}
\end{figure}

The Markovian structure of the latent states in the models used provides additional information that can be used in the structure of the vine. Denoting the number of elements in $\boldsymbol{\theta}$ with $k$, allow the first $k$ trees to contain edges that correspond to copulas between different elements of $\boldsymbol{\theta}$, and an element of $\boldsymbol{\theta}$ and $x_t$. The $k+1'th$ tree edges correspond to the copulas $c(x_0, x_1 | \boldsymbol{\theta}, c(x_1, x_2 | \boldsymbol{\theta}), \dots, c(x_{T-1}, x_T | \boldsymbol{\theta}$ and each further tree contains only bivariate copulas between $x_t$ and $x_{t+j}$ conditioned on $x_{t+l}$, where $j > l \geq 1$. Markovian latent states implies that $x_t$ and $x_{t+2}$ are conditionally independent given $x_{t+1}$, so each copula in tree $k+2$ onwards can be set to the independence copula, reducing the number of copula parameters in the vine to be linear in $T$ instead of quadratic. To allow the vine structure to be easily extended to new latent variables $x_{T+1:T+S}$, each $x_t$ node should have edges to the same $\boldsymbol{\theta}$ node. Di{\ss}mann's algorith (Di{\ss}mann, Brechmann, Czado and Kurowicka 2013) can be used to select a well fitting vine structure subject to these restrctions.

@@

We transform $\sigma^2_y$ and $\sigma^2_x$ so that $\boldsymbol{\theta} = \{\gamma, \phi, \ln(\sigma^2_y), \ln(\sigma^2_x) \} \in \mathbb{R}^4$ and consider three different functional forms for the approximating distribution $q(\boldsymbol{\theta}, x_{1:T} | \boldsymbol{\lambda})$:
\begin{enumerate}
\item $q(\boldsymbol{\theta}, x | \boldsymbol{\lambda}) = \mathcal{N}(\boldsymbol{\psi}, \textbf{V})$, where $\textbf{V}$ is a diagonal matrix.
\item $q(\boldsymbol{\theta}, x | \boldsymbol{\lambda}) = \mathcal{N}(\boldsymbol{\psi}, \boldsymbol{\Sigma})$, where $\boldsymbol{\Sigma}$ allows non-zero off-diagonal terms.
\item  $q(\boldsymbol{\theta}, x | \boldsymbol{\lambda}) = \prod \left(q(\theta_i) q(x_{it})\right) c(\boldsymbol{\theta}, x)$ where the form of the copula $c$ and marginals are informed by an out of date MCMC sample.
\end{enumerate}
The first two functional forms are transformation of standard normal noise variables while the third is a transformation of uniform noise variables.

We use three MCMC based approaches for forecasting, listed in increasing computational complexity:
\begin{enumerate}
\item Run the full MCMC to obtain $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})$ and perform an $S+h$ step ahead forecast of $p(y_{T+S+h} | y_{1:T})$, marginalising out the variables $\boldsymbol{\theta}$ and $x_{1:T}$ using draws from $p(\boldsymbol{\theta}, x_{1:T} | y_{1:T})$.
\item Using the MCMC from approach 1, use the Kalman filter to update the distribution of $x_{T+1:T+S+h}$ after $y_{T+1:T+S}$ is observed to obtain the $h$ step ahead forecast of $p(y_{T+S+h} | y_{1:T+S})$. This method does not condition the posterior distribution of $\boldsymbol{\theta}$ on new observations.
\item Run the full MCMC to obtain $p(\boldsymbol{\theta}, x_{1:T+S} | y_{1:T+S})$ and perform a $h$ step ahead forecast of $p(y_{T+S+h} | y_{1:T+S})$, marginalising out the variables $\boldsymbol{\theta}$ and $x_{1:T+S}$. We assume that the computational time to run this MCMC after observation of $y_{T+S}$ is greater than the time period between $y_{T+S}$ and $y_{T+S+h}$, but this provides a benchmark for comparision with Variational Bayes methods.
\end{enumerate}
